Manuscript Figures for Genome Biology

### 2019-03-28 manuscript_figures/manuscript_figures_main.Rmd
Response to reviewer 1, comment 1
Add iCOBRA-style plots to combine Figures 2A-3A, and Figures 2B-3B.

## exclude 'enigma-so4-otus' results with no significant calls
objdat <- dplyr::filter(objdat, casestudy != "enigma-so4-otus")

## create table of case study-to-assay mapping
studies <- tibble(
    casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "baxter-genus", 
                  "goodrich-genus", "papa-genus", "schubert-otus", "brain", "mir200c",
                  "human-mast-det", "human-mast-mean", "human-scdd-det",
                  "human-scdd-mean", "human-wilcox-det", "human-wilcox-mean",
                  "mouse-mast-det", "mouse-mast-mean", "mouse-scdd-det",
                  "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 7),
              rep("RNA-seq", 2), rep("scRNA-seq", 12)))

## add assay labels to results table
objdat <- left_join(objdat, studies, by = "casestudy")
objdat <- dplyr::filter(objdat, !is.na(assay))

@@

p4b <- ggplot(objdat_joint, aes(x = casestudy, y = method, fill = propMaxRej)) +
    scale_x_discrete("Case Study", expand = c(0, 0)) +
    scale_y_discrete("Method", expand = c(0, 0)) +
    geom_tile() +
    geom_text(aes(label = lab), data = filter(objdat_props, method %in% methodset),
              size = 1.8, color = "white") +
    geom_text(aes(label = should), size = 5, color = "black") +
    geom_tile(data = filter(objdat_joint, plotrow == "fake"), fill = "white") + 
    geom_text(aes(label = cnt), data = dplyr::filter(objdat_joint, !is.na(cnt)), size = 2) + 
    scale_fill_distiller("% Rejected\n(relative to max)",
                         palette = "Blues", direction = 1, limits = c(0, 1),
                         labels = scales::percent) +
    facet_grid(plotrow ~ assay, scales = "free", space = "free") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
          strip.text.y = element_blank(),
          legend.position = c(0,-0.32), legend.justification = c(0, 1),
          legend.direction = "horizontal",
          legend.text = element_text(size = 7),
          legend.title = element_text(size = 9),
          legend.margin = margin(6, 10, 6, 10))
Titlep4b <- ggdraw() +
    draw_label(expression(paste(bold("Number of rejections in case studies"))))
p4b <- plot_grid(Titlep4b, p4b, nrow = 2, rel_heights = c(0.075,1), 
                 labels = c("B", ""))

@@

>>>>

## exclude 'enigma-so4-otus' results with no significant calls
objdat <- dplyr::filter(objdat, !grepl("so4", casestudy))

## create table of case study-to-assay mapping
studies <- tibble(
    casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "baxter-genus", 
                  "goodrich-genus", "papa-genus", "schubert-otus",
                  "enigma-al-otus-abun", "enigma-ph-otus-abun", "enigma-so4-otus-abun",
                  "baxter-genus-mean-abun", "goodrich-genus-abun", "papa-genus-abun",
                  "schubert-otus-abun",
                  "brain", "mir200c",
                  "human-mast-det", "human-mast-mean", "human-scdd-det",
                  "human-scdd-mean", "human-wilcox-det", "human-wilcox-mean",
                  "mouse-mast-det", "mouse-mast-mean", "mouse-scdd-det",
                  "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 14),
              rep("RNA-seq", 2), rep("scRNA-seq", 12)))

## add assay labels to results table
objdat <- left_join(objdat, studies, by = "casestudy")
objdat <- dplyr::filter(objdat, !is.na(assay == "Microbiome"))

# add covariate label to microbiome ubiquity 
objdat <- objdat %>%
          mutate(casestudy = ifelse(assay == "Microbiome", 
                                    ifelse(!grepl("abun", casestudy),
                                           paste0(casestudy, "-ubiq"),
                                           casestudy),
                                    casestudy)) %>%
          mutate(casestudy = gsub("mean-abun", "abun", casestudy))

@@

p4b <- ggplot(objdat_joint, aes(x = casestudy, y = method, fill = propMaxRej)) +
    scale_x_discrete("Case Study", expand = c(0, 0)) +
    scale_y_discrete("Method", expand = c(0, 0)) +
    geom_tile() +
    geom_text(aes(label = lab), data = filter(objdat_props, method %in% methodset),
              size = 1.8, color = "white") +
    geom_text(aes(label = should), size = 5, color = "black") +
    geom_tile(data = filter(objdat_joint, plotrow == "fake"), fill = "white") + 
    geom_text(aes(label = cnt), data = dplyr::filter(objdat_joint, !is.na(cnt)), size = 2) + 
    scale_fill_distiller("% Rejected\n(relative to max)",
                         palette = "Blues", direction = 1, limits = c(0, 1),
                         labels = scales::percent) +
    facet_grid(plotrow ~ assay, scales = "free", space = "free") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
          strip.text.y = element_blank(),
          legend.position = c(0,-0.38), legend.justification = c(0, 1),
          legend.direction = "horizontal",
          legend.text = element_text(size = 7),
          legend.title = element_text(size = 9),
          legend.margin = margin(6, 10, 6, 10))
Titlep4b <- ggdraw() +
    draw_label(expression(paste(bold("Number of rejections in case studies"))))
p4b <- plot_grid(Titlep4b, p4b, nrow = 2, rel_heights = c(0.075,1), 
                 labels = c("B", ""))

@@

# Revisions - Reviewer 1, comment 1:

This is a figure that combines Figure 2A and 3A into one plot (mean FDR vs mean TPR), with open or
closed circles representing whether or not the nominal error rate is controlled (on average) at three 
points (0.01, 0.05, 0.10).

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
r1.1a <- plotsim_average(res5d, met=c("FDR", "TPR"), filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) +
      theme(legend.position="none") +
      ggtitle("FDR vs TPR") +
      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))

res5d.sum <- res5d %>% 
  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
  summarize(n = sum(!is.na(value)),
            se = sd(value, na.rm = TRUE) / sqrt(n),
            value = mean(value, na.rm = TRUE)) %>%
  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
         performanceMetric %in% c("FDR", "TPR"),
         !blabel %in% excludeSet,
         is.na(param.alpha) | (param.alpha == alpha))

tsba_m <- NULL 
met = c("FDR", "TPR")
for (m in seq_along(met)){
     tmp <- res5d.sum %>% 
     filter(performanceMetric == met[m]) %>%
     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
     if (m > 1){
      tsba_m <- left_join(tsba_m, tmp, 
                          by = c("Method", "alpha", "n", 
                                 "param.alpha", "blabel"))
  }else{
      tsba_m <- tmp
  }
}
res5d.sum <- tsba_m
res5d.sum <- res5d.sum %>%
  mutate(control = value.x < alpha) %>%
  mutate(col2 = ifelse(control, col, "white"))

# add color palette
res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")

r1.1a <- r1.1a +
  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
             shape = control), size = 3.5) +
  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
             shape = 19, size = 2.5, color = "white") +
  scale_shape_manual(values = c(1, 19)) + 
  ggtitle("In silico experiments") +
  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
                         breaks=seq(0, 1, by=0.02))
```

Next we'll do the same for 2B and 3B (polyester).

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
r1.1b <- plotsim_average(res5dp, met=c("FDR", "TPR"), filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) +
      theme(legend.position="none") +
      ggtitle("FDR vs TPR") +
      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))

res5d.sum <- res5dp %>% 
  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
  summarize(n = sum(!is.na(value)),
            se = sd(value, na.rm = TRUE) / sqrt(n),
            value = mean(value, na.rm = TRUE)) %>%
  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
         performanceMetric %in% c("FDR", "TPR"),
         !blabel %in% excludeSet,
         is.na(param.alpha) | (param.alpha == alpha))

tsba_m <- NULL 
met = c("FDR", "TPR")
for (m in seq_along(met)){
     tmp <- res5d.sum %>% 
     filter(performanceMetric == met[m]) %>%
     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
     if (m > 1){
      tsba_m <- left_join(tsba_m, tmp, 
                          by = c("Method", "alpha", "n", 
                                 "param.alpha", "blabel"))
  }else{
      tsba_m <- tmp
  }
}
res5d.sum <- tsba_m
res5d.sum <- res5d.sum %>%
  mutate(control = value.x < alpha) %>%
  mutate(col2 = ifelse(control, col, "white"))

# add color palette
res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")

r1.1b <- r1.1b +
  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
             shape = control), size = 3.5) +
  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
             shape = 19, size = 2.5, color = "white") +
  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
             shape = 1, size = 3.5) +
  scale_shape_manual(values = c(1, 19)) +
  ggtitle("Polyester") +
  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
                         breaks=seq(0, 1, by=0.02))
```

Put them together and save.

```{r}
r1.1 <- plot_grid(r1.1a, r1.1b, rel_widths = c(1,1.05))
r1.1 <- plot_grid(r1.1, get_legend(p2a), rel_widths = c(1, .2)) 
r1.1
ggsave(file.path(outdir, "Figure_R1.1.pdf"), width = 8, height = 3)
```
